---
title: "Data Validation Report"
subtitle: "Scottish Pupils Administrative Dataset"
date: "`r Sys.Date()`"
format:
  html:
    theme: cosmo
    toc: true
    toc-title: "Contents"
    code-fold: true
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
---
```{r setup}
library(tidyverse)
library(targets)
```

## Overview

This report summarises the results of automated validation checks run 
against the Scottish pupils administrative dataset. Each check identifies 
a specific category of data quality issue.
```{r load-results}
tar_load(validation_results)
tar_load(validation_summary)
```

## Summary
```{r summary-table}
validation_summary |>
  mutate(
    status = if_else(status == "PASS",
                     "✅ PASS",
                     "❌ FAIL")
  ) |>
  rename(
    `Check`    = check_name,
    `Status`   = status,
    `Issues`   = n_issues
  ) |>
  knitr::kable()
```
```{r overall-status}
n_failed <- validation_summary |> filter(status == "FAIL") |> nrow()
n_passed <- validation_summary |> filter(status == "PASS") |> nrow()
total    <- nrow(validation_summary)
```

Out of **`r total`** checks run, **`r n_passed`** passed and 
**`r n_failed`** failed.

## Check Details

### Duplicate Pupil IDs
```{r duplicates}
validation_results$details[[1]]$detail |>
  knitr::kable()
```

### Missing Required Fields
```{r missing-summary}
validation_results$details[[2]]$missing_by_field |>
  knitr::kable()
```

### Invalid Gender Codes
```{r gender}
validation_results$details[[3]]$detail |>
  select(pupil_id, gender) |>
  knitr::kable()
```

### Attendance Out of Range
```{r attendance}
validation_results$details[[4]]$detail |>
  select(pupil_id, attendance_rate) |>
  knitr::kable()
```

### Stage and School Type Mismatch
```{r stage}
validation_results$details[[5]]$detail |>
  select(pupil_id, school_type, stage) |>
  knitr::kable()
```

### Future or Implausible Dates
```{r dates}
validation_results$details[[6]]$detail |>
  select(pupil_id, date_of_birth, date_enrolled) |>
  knitr::kable()
```